Ancestry linkage in hybrid zones

### 2016-02-14 drafts/hybrid_zone.tex

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}. When formerly isolated populations come back into contact \yb{the exchange of genes across a geographic hybrid zone can establish a gradient of ancestry across geography, \citep{Barton1985}}. In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD). If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. \yb{It is clear that selection against admixed ancestry at one locus will influence patterns of ancestry at physically linked loci, however the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation. The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. \alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. }

While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. 
These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

% Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci.  %YB Genome-wide sounds wrong to me. is this ok?
Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci. 
%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.   % YB I don't think we are actually asking about density of selected sites, are we? 
Our goal here is  to understand how strength of \yb{selection, recombination} rate, and hybrid zone age interact to produce observed patterns of ancestry in present day populations.  
This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions. In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact. In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.

Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center. This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events. Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.

>>>>

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
\yb{Thus, ancestry originating from the two diverged populations spreads from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}. 
%In addition to steep clines in ancestry,
\yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}  
\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both of the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} 
However, subsets of the genome that are under selection can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. 
\yb{However the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher; Chapman \& Thomson; Barid et al)}.   
The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection \yb{(CITE CLINES)}. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. 
\alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
These models predict cline width to be narrowest \yb{(i.e. the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider. %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci. 
%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.   % YB I don't think we are actually asking about density of selected sites, are we? 
Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}. %in present day populations.  
This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness. 
In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \yb{(CITE: hapmix, globetrotter)}. 
\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci (CITE: clinefit, gbc), but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}  
%In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.

Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center. 
This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events. 
Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.  
\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}
